Clustering of Distributions on 1-Dimensional Manifolds

### 2021-09-29 scratch/distributions-on-1d-manifolds.Rmd
syncing stuff

---
title: |
  | Clustering of Distributions on 1-Dimensional Manifolds
output:
  pdf_document:
    keep_tex: true
    citation_package: natbib
fontsize: 11pt
# geometry: "left=1in,right=1in,top=1in,bottom=1in"
urlcolor: blue
header-includes:
- \usepackage{setspace}
- \usepackage{float}
- \usepackage{mathtools}
- \usepackage{natbib}
- \usepackage[linesnumbered,ruled,vlined]{algorithm2e} 
- \setcitestyle{numbers,square,comma}
- \usepackage{verbatim}
- \usepackage{amsthm}
- \usepackage{comment}
bibliography: proposal.bib
---

\newcommand{\diag}{\mathrm{diag}}
\newcommand{\tr}{\mathrm{Tr}}
\newcommand{\blockdiag}{\mathrm{blockdiag}}
\newcommand{\indep}{\stackrel{\mathrm{ind}}{\sim}}
\newcommand{\iid}{\stackrel{\mathrm{iid}}{\sim}}
\newcommand{\Bernoulli}{\mathrm{Bernoulli}}
\newcommand{\Betadist}{\mathrm{Beta}}
\newcommand{\Uniform}{\mathrm{Uniform}}
\newcommand{\BG}{\mathrm{BernoulliGraph}}
\newcommand{\Categorical}{\mathrm{Categorical}}
\newtheorem{definition}{Definition}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\theoremstyle{example}
\newtheorem*{example}{Example}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\as}{\stackrel{\mathrm{a.s.}}{\to}}

@@

Suppose we have two manifolds $\mathcal{M}_1$ and $\mathcal{M}_2$ $\in \mathbb{R}^d$, each of length 1, defined by $f_1(t)$ and $f_2(t)$ respectively ($f_i : [0, 1] \mapsto \mathbb{R}^d$). Define $\delta$ as the minimum distance between the two manifolds, i.e., $\delta = \min_t \|f_1(t) - f_2(t)\|$, and let $\delta > 0$. For now, restrict each $f_i$ such that the distance along the manifold between $f_i(t)$ and $f_i(s)$ is equal to the difference between $t$ and $s$ (this also implies that each manifold is of length 1). 
We sample $T_1, ..., T_n \stackrel{iid}{\sim} F$ for continuous $F$ with support $[0, 1]$ and use $f_1$ to map the first $n_1$ points to $\mathcal{M}_1$ and $f_2$ to map the remaining $n_2 = n - n_1$ points to $\mathcal{M}_2$. Let $X_i = f_1(T_i)$ and $Y_j = f_2(T_j)$.
Without loss of generality, assume $n_1 \leq n_2$.

@@

Thus if we are able to show that Eq. (\ref{eq:cdf}) converges to 1, we are done.

@@

\begin{theorem}
Let $X_1, ..., X_n \iid \Uniform(0, 1)$. 
Then for any $\delta > 0$, $P(\max_i X_{(i+1)} - X_{(i)} < \delta) \to 1$ as 
$n \to \infty$.
\end{theorem}

@@

\begin{theorem}
Let $X_1, ..., X_n \iid F$ with support $[0, 1]$, and suppose $f(x)$ is continuous and $f(x) \geq a > 0$ everywhere on the support. 
Let $D_i = X_{(i+1)} - X_{(i)}$. 
Then for any $\epsilon > 0$, there exists $N \in \mathbb{N}$ such that $P(\max_i D_i < \delta) > 1 - \epsilon$ when $n > N$.
\end{theorem}

@@

$$
\begin{aligned}
P(\max_i D_i < \delta) & \geq (P(D_i < \delta))^{n-1} \\
& \approx (1 - (1 - a \delta)^n + (1 - F(1-\delta) - a \delta)^n)^{n-1} \\
& \to 1
\end{aligned}
$$

\end{proof}

@@

>>>>

---
title: |
  | Clustering of Distributions on 1-Dimensional Manifolds
output:
  pdf_document:
    keep_tex: true
    citation_package: natbib
fontsize: 11pt
# geometry: "left=1in,right=1in,top=1in,bottom=1in"
urlcolor: blue
header-includes:
- \usepackage{setspace}
- \usepackage{float}
- \usepackage{mathtools}
- \usepackage{natbib}
- \usepackage[linesnumbered,ruled,vlined]{algorithm2e} 
- \setcitestyle{numbers,square,comma}
- \usepackage{verbatim}
- \usepackage{amsthm}
- \usepackage{comment}
bibliography: misc.bib
---

\newcommand{\diag}{\mathrm{diag}}
\newcommand{\tr}{\mathrm{Tr}}
\newcommand{\blockdiag}{\mathrm{blockdiag}}
\newcommand{\indep}{\stackrel{\mathrm{ind}}{\sim}}
\newcommand{\iid}{\stackrel{\mathrm{iid}}{\sim}}
\newcommand{\Bernoulli}{\mathrm{Bernoulli}}
\newcommand{\Betadist}{\mathrm{Beta}}
\newcommand{\Uniform}{\mathrm{Uniform}}
\newcommand{\BG}{\mathrm{BernoulliGraph}}
\newcommand{\Categorical}{\mathrm{Categorical}}
\newcommand{\Multinomial}{\mathrm{Multinomial}}
\newcommand{\RDPG}{\mathrm{RDPG}}
\newcommand{\GRDPG}{\mathrm{GRDPG}}
\newtheorem{definition}{Definition}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\theoremstyle{example}
\newtheorem*{example}{Example}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\as}{\stackrel{\mathrm{a.s.}}{\to}}

@@

Suppose we have two manifolds $\mathcal{M}_1$ and $\mathcal{M}_2$ $\in \mathbb{R}^d$, each of length 1, defined by $f_1(t)$ and $f_2(t)$ respectively ($f_k : [0, 1] \mapsto \mathbb{R}^d$). 
Define $\delta$ as the minimum distance between any two manifolds, i.e., $\delta = \max_{k, \ell} \min_{s, t} \|f_k(s) - f_\ell(t)\|$, and let $\delta > 0$. 
Restrict each $f_k$ such that the distance along the manifold between $f_k(t)$ and $f_k(s)$ is equal to the difference between $t$ and $s$, i.e., $f_k$ is the arclength parameterization of $\mathcal{M}_k$ (this also implies that each manifold is of length 1). 
Labels are sampled as $Z_1, ..., Z_n \iid \Multinomial(\alpha_1, \alpha_2)$, and we define $n_k$ as the number of labels with label $k$. 
Sample $T_1, ..., T_n \iid F$ for some distribution $F$ with support $[0, 1]$. 
Finally, let $X_i = f_{Z_i}(T_i)$ for each $i \in [n]$ be the latent vector of vertex $i$, and gather the latent vectors as $X = \begin{bmatrix} X_1 & \cdots & X_n \end{bmatrix}$. 
We observe $A \sim \RDPG(X)$ (or $A \sim \GRDPG_{p,q}(X)$) and wish to recover $Z_1, ..., Z_n$.

@@

@@

\begin{theorem}
Let $X_1, ..., X_n \iid \Uniform(0, 1)$. 
Then for any $\epsilon$ and $\delta > 0$, 
there exists an $N = O \big(\frac{-\log \epsilon}{\delta} \big)$ such that 
$P(\max_i X_{(i+1)} - X_{(i)} < \delta) \geq 1 - \epsilon$ when $n > N$.
\end{theorem}

@@

\begin{theorem}
Let $X_1, ..., X_n \iid F$ with support $[0, 1]$, and suppose $f(x)$ is continuous and $f(x) \geq a > 0$ everywhere on the support. 
Let $D_i = X_{(i+1)} - X_{(i)}$. 
Then for any $\epsilon > 0$, there exists $N \in \mathbb{N}$ such that $P(\max_i D_i < \delta) \geq 1 - \epsilon$ when $n > N$.
\end{theorem}

@@

$$
\begin{aligned}
P(\max_i D_i < \delta) & = P(\text{all } D_i < \delta) \\
& = 1 - P(\text{some } D_i > \delta) \\
& \geq 1 - \sum_i^{n-1} P(D_i > \delta) \\
& = 1 + (n - 1) (1 - a \delta)^n - (n - 1) (1 - F(1 - \delta) - a \delta)^n
\end{aligned}
$$

This converges to $1$ in the limit $n \to \infty$.
\end{proof}

## Extension to multidimensional manifolds

Here, we extend the results on the line to the unit hypercube. The following theorem is a direct consequence of Lemma 2 of \citet{trosset2020rehabilitating}.

\begin{theorem}
Let $X_1, ..., X_n \iid F$ with support $[0, 1]^r$, and $f(x) \geq a > 0$ everywhere on the support. Define $E_n$ as the event that every point on the support is within distance $\delta / 2$ to $X_i$ for $i \in [n]$. Then for any $\epsilon > 0$, there exists $N = O(-\log \epsilon)$ such that $P(E_n) > 1 - \epsilon$ when $n \geq N$.
\end{theorem}

@@

# References

### 2021-09-29 scratch/distributions-on-1d-manifolds.tex
syncing stuff

\newcommand{\diag}{\mathrm{diag}}
\newcommand{\tr}{\mathrm{Tr}}
\newcommand{\blockdiag}{\mathrm{blockdiag}}
\newcommand{\indep}{\stackrel{\mathrm{ind}}{\sim}}
\newcommand{\iid}{\stackrel{\mathrm{iid}}{\sim}}
\newcommand{\Bernoulli}{\mathrm{Bernoulli}}
\newcommand{\Betadist}{\mathrm{Beta}}
\newcommand{\Uniform}{\mathrm{Uniform}}
\newcommand{\BG}{\mathrm{BernoulliGraph}}
\newcommand{\Categorical}{\mathrm{Categorical}}
\newtheorem{definition}{Definition}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\theoremstyle{example}
\newtheorem*{example}{Example}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\as}{\stackrel{\mathrm{a.s.}}{\to}}

@@

Suppose we have two manifolds \(\mathcal{M}_1\) and \(\mathcal{M}_2\)
\(\in \mathbb{R}^d\), each of length 1, defined by \(f_1(t)\) and
\(f_2(t)\) respectively (\(f_i : [0, 1] \mapsto \mathbb{R}^d\)). Define
\(\delta\) as the minimum distance between the two manifolds, i.e.,
\(\delta = \min_t \|f_1(t) - f_2(t)\|\), and let \(\delta > 0\). For
now, restrict each \(f_i\) such that the distance along the manifold
between \(f_i(t)\) and \(f_i(s)\) is equal to the difference between
\(t\) and \(s\) (this also implies that each manifold is of length 1).
We sample \(T_1, ..., T_n \stackrel{iid}{\sim} F\) for continuous \(F\)
with support \([0, 1]\) and use \(f_1\) to map the first \(n_1\) points
to \(\mathcal{M}_1\) and \(f_2\) to map the remaining \(n_2 = n - n_1\)
points to \(\mathcal{M}_2\). Let \(X_i = f_1(T_i)\) and
\(Y_j = f_2(T_j)\). Without loss of generality, assume \(n_1 \leq n_2\).

@@

\begin{theorem}
Let $X_1, ..., X_n \stackrel{\mathrm{iid}}{\sim}\mathrm{Uniform}(0, 1)$. 
Then for any $\delta > 0$, $P(\max_i X_{(i+1)} - X_{(i)} < \delta) \to 1$ as 
$n \to \infty$.
\end{theorem}

@@

\begin{theorem}
Let $X_1, ..., X_n \stackrel{\mathrm{iid}}{\sim}F$ with support $[0, 1]$, and suppose $f(x)$ is continuous and $f(x) \geq a > 0$ everywhere on the support. 
Let $D_i = X_{(i+1)} - X_{(i)}$. 
Then for any $\epsilon > 0$, there exists $N \in \mathbb{N}$ such that $P(\max_i D_i < \delta) > 1 - \epsilon$ when $n > N$.
\end{theorem}

@@

$$
\begin{aligned}
P(\max_i D_i < \delta) & \geq (P(D_i < \delta))^{n-1} \\
& \approx (1 - (1 - a \delta)^n + (1 - F(1-\delta) - a \delta)^n)^{n-1} \\
& \to 1
\end{aligned}
$$

\end{proof}

@@

  \bibliography{proposal.bib}

>>>>

\newcommand{\diag}{\mathrm{diag}}
\newcommand{\tr}{\mathrm{Tr}}
\newcommand{\blockdiag}{\mathrm{blockdiag}}
\newcommand{\indep}{\stackrel{\mathrm{ind}}{\sim}}
\newcommand{\iid}{\stackrel{\mathrm{iid}}{\sim}}
\newcommand{\Bernoulli}{\mathrm{Bernoulli}}
\newcommand{\Betadist}{\mathrm{Beta}}
\newcommand{\Uniform}{\mathrm{Uniform}}
\newcommand{\BG}{\mathrm{BernoulliGraph}}
\newcommand{\Categorical}{\mathrm{Categorical}}
\newcommand{\Multinomial}{\mathrm{Multinomial}}
\newcommand{\RDPG}{\mathrm{RDPG}}
\newcommand{\GRDPG}{\mathrm{GRDPG}}
\newtheorem{definition}{Definition}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\theoremstyle{example}
\newtheorem*{example}{Example}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\as}{\stackrel{\mathrm{a.s.}}{\to}}

@@

Suppose we have two manifolds \(\mathcal{M}_1\) and \(\mathcal{M}_2\)
\(\in \mathbb{R}^d\), each of length 1, defined by \(f_1(t)\) and
\(f_2(t)\) respectively (\(f_k : [0, 1] \mapsto \mathbb{R}^d\)). Define
\(\delta\) as the minimum distance between any two manifolds, i.e.,
\(\delta = \max_{k, \ell} \min_{s, t} \|f_k(s) - f_\ell(t)\|\), and let
\(\delta > 0\). Restrict each \(f_k\) such that the distance along the
manifold between \(f_k(t)\) and \(f_k(s)\) is equal to the difference
between \(t\) and \(s\), i.e., \(f_k\) is the arclength parameterization
of \(\mathcal{M}_k\) (this also implies that each manifold is of length
1). Labels are sampled as
\(Z_1, ..., Z_n \stackrel{\mathrm{iid}}{\sim}\mathrm{Multinomial}(\alpha_1, \alpha_2)\),
and we define \(n_k\) as the number of labels with label \(k\). Sample
\(T_1, ..., T_n \stackrel{\mathrm{iid}}{\sim}F\) for some distribution
\(F\) with support \([0, 1]\). Finally, let \(X_i = f_{Z_i}(T_i)\) for
each \(i \in [n]\) be the latent vector of vertex \(i\), and gather the
latent vectors as
\(X = \begin{bmatrix} X_1 & \cdots & X_n \end{bmatrix}\). We observe
\(A \sim \mathrm{RDPG}(X)\) (or \(A \sim \mathrm{GRDPG}_{p,q}(X)\)) and
wish to recover \(Z_1, ..., Z_n\).

@@

\begin{theorem}
Let $X_1, ..., X_n \stackrel{\mathrm{iid}}{\sim}\mathrm{Uniform}(0, 1)$. 
Then for any $\epsilon$ and $\delta > 0$, 
there exists an $N = O \big(\frac{-\log \epsilon}{\delta} \big)$ such that 
$P(\max_i X_{(i+1)} - X_{(i)} < \delta) \geq 1 - \epsilon$ when $n > N$.
\end{theorem}

@@

\begin{theorem}
Let $X_1, ..., X_n \stackrel{\mathrm{iid}}{\sim}F$ with support $[0, 1]$, and suppose $f(x)$ is continuous and $f(x) \geq a > 0$ everywhere on the support. 
Let $D_i = X_{(i+1)} - X_{(i)}$. 
Then for any $\epsilon > 0$, there exists $N \in \mathbb{N}$ such that $P(\max_i D_i < \delta) \geq 1 - \epsilon$ when $n > N$.
\end{theorem}

@@

$$
\begin{aligned}
P(\max_i D_i < \delta) & = P(\text{all } D_i < \delta) \\
& = 1 - P(\text{some } D_i > \delta) \\
& \geq 1 - \sum_i^{n-1} P(D_i > \delta) \\
& = 1 + (n - 1) (1 - a \delta)^n - (n - 1) (1 - F(1 - \delta) - a \delta)^n
\end{aligned}
$$

This converges to $1$ in the limit $n \to \infty$.
\end{proof}

\hypertarget{extension-to-multidimensional-manifolds}{%
\subsection{Extension to multidimensional
manifolds}\label{extension-to-multidimensional-manifolds}}

Here, we extend the results on the line to the unit hypercube. The
following theorem is a direct consequence of Lemma 2 of
\citet{trosset2020rehabilitating}.

\begin{theorem}
Let $X_1, ..., X_n \stackrel{\mathrm{iid}}{\sim}F$ with support $[0, 1]^r$, and $f(x) \geq a > 0$ everywhere on the support. Define $E_n$ as the event that every point on the support is within distance $\delta / 2$ to $X_i$ for $i \in [n]$. Then for any $\epsilon > 0$, there exists $N = O(-\log \epsilon)$ such that $P(E_n) > 1 - \epsilon$ when $n \geq N$.
\end{theorem}

@@

\renewcommand\refname{References}
  \bibliography{misc.bib}
